---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rgridstatus

<!-- badges: start -->
[![R-CMD-check](https://github.com/andypicke/rgridstatus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andypicke/rgridstatus/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of rgridstatus is to provide a R API wrapper for the [GridStatusIO](https://www.gridstatus.io/home) [API](https://www.gridstatus.io/api).

## Installation

You can install the development version of rgridstatus from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andypicke/rgridstatus")
```

You will need to register for an API key. By default, all functions assume you have stored your API key in your *.Renviron* file with the name *GRIDSTATUS_API_KEY*.

## Examples

### Get info on available datasets:

```{r example}
library(rgridstatus)
info <- get_available_datasets()
head(info)
```

### Get list of dataset updates

```{r}

updates <- get_dataset_updates()
head(updates)

```

### Download a dataset

```{r}

df <- get_gridstatus_dataset("caiso_fuel_mix")

str(df)

```

