---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rgridstatus

<!-- badges: start -->

[![R-CMD-check](https://github.com/andypicke/rgridstatus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/andypicke/rgridstatus/actions/workflows/R-CMD-check.yaml) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![Project Status: WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

<!-- badges: end -->

The goal of rgridstatus is to provide a R API wrapper for the [GridStatusIO](https://www.gridstatus.io/home) [API](https://www.gridstatus.io/api).

See also my [blog post](https://andypicke.quarto.pub/portfolio/posts/rGridStatus/rGridStatus.html) describing some features of the package.

## Installation

You can install the development version of [rgridstatus](https://github.com/andypicke/rgridstatus) from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andypicke/rgridstatus")
```

You will need to register for an API key. By default, all functions assume you have stored your API key in your *.Renviron* file with the name *GRIDSTATUS_API_KEY*.

## Examples

### Get info on datasets available from GridStatus API:

```{r example}

library(rgridstatus)
library(tidyverse)

info <- get_available_datasets()
head(info)
```

### Get unique data source names

Most of the "sources" are [ISOs](https://en.wikipedia.org/wiki/Regional_transmission_organization_(North_America)).

```{r}

# input is info from get_available_datasets()
sources <- get_source_names(info)
sources

```

### Get dataset names for one operator/source

The dataset names all start with the ISO/source, so it is easy to filter by those:

```{r}

info_caiso <- get_available_datasets() |> 
  dplyr::filter(stringr::str_detect(name,"^CAISO_")) # name starts with "CAISO_"

```

### Get a list of dataset updates

```{r}

updates <- get_dataset_updates()
head(updates)

```

Again, you can filter the results to get updates for a specific ISO/source:
```{r}

updates_caiso <- updates |> dplyr::filter(stringr::str_detect(dataset, "^caiso_"))
updates_caiso
```


### Download a dataset

```{r}

df <- get_gridstatus_dataset(wh_dataset = "caiso_fuel_mix", 
                             start_time = "2024-09-03", 
                             end_time = "2024-09-05")

str(df)

```


### Pivot fuel mix dataframe to long format for easier plotting

{rgridstatus} contains a convenience function *pivot_gen_long()* to pivot the fuel mix dataframe to a long format. This can make it easier to plot multiple timeseries.

```{r}

df_long <- pivot_gen_long(df)

head(df_long)

```

With the data in a long format, we can just specify that the color or linetype should correspond to the *fuel_type*.

```{r}


df_long |>
  ggplot(aes(datetime_local, generation)) +
  geom_line(aes(color = fuel_type)) 


```

